---
title: 'R Notebook: data-intro'
author: "Peyton Urquhart [peyton.urquhart]"
output:
  html_document:
    df_print: paged
---

```{r, echo=FALSE}
require("ggplot2")
require("reticulate")
#install.packages("C:/_git_/humanVerseWSU/humanVerseWSU_0.1.4.zip", repos=NULL, type="source");
library(humanVerseWSU);
#install.packages("C:/_git_/monte-imdb/imdb_0.1.1.zip", repos=NULL, type="source");
library(imdb);
```

#### Search machine for Python

I couldn't get some Python packages to load.. Need to read up on virtualenv. I
will keep the initial setup to try for the next assignment.

```{r, echo=TRUE}
reticulate::py_config();
```

#### Was Python Found ?

```{r, echo=FALSE}
if (reticulate::py_available() == FALSE)
{
  print("ERROR: python installation not found");
} else {
  print("SUCCESS: python installation found");
}
```

```{r, echo=FALSE}
github.local = "C:/_git_/peyton.urquhart/";
github.remote = "https://raw.githubusercontent.com/peytonurquhart/WSU_STAT_419/main/";
script = "functions/functions-imdb.R";
```

#### Local or Remote ?

```{r}
which.github = "local";
#which.github = "remote";
```

```{r, echo=FALSE}
if(which.github == "remote")
{
  include.me = paste0( github.remote, script);
  library(devtools)
  library(rstudioapi)
  source_url(include.me);
  
} else if (which.github == "local") {
  
    include.me = paste0( github.local, script);
    source(include.me);
}
```

```{r, chunk-setup, include=FALSE, echo=FALSE}
# include=FALSE means it is running, but invisible ...
knitr::opts_chunk$set(echo = TRUE);
knitr::opts_chunk$set(warning = FALSE);
knitr::opts_chunk$set(message = FALSE);
```

```{r, chunck-loadData, include=TRUE}
loadDataIMDB();
```

## Will Smith

Build a dataframe "will.df" for Will Smith with 2 primary sections:

will.df$info containing a single row of information about Will Smith.

will.df$movies containing a large dataframe where each row represents one of Will
Smiths movies.

```{r}
will.df <- list(movies = IMDB.getMoviesForPerson("nm0000226"), info = IMDB.getPersonInfo("nm0000226"));
will.df$movies;
will.df$info;
```

## Denzel Washington

Build a dataframe "denzel.df" for Denzel Washington with 2 primary sections:

denzel.df$info containing a single row of information about Denzel.

denzel.df$movies containing a large dataframe where each row represents one of
Denzel Washingtons movies.

```{r}
denzel.df <- list(movies = IMDB.getMoviesForPerson("nm0000243"), info = IMDB.getPersonInfo("nm0000243"));
denzel.df$movies;
denzel.df$info;
```

## Common Movies

Search the dataframes for matching ttid(s). Build a new dataframe for each match
resulting in a dataframe of movies starring both Will and Denzel

```{r}
findCommonMovies(will.df$movies, denzel.df$movies);
```

## Summary

I wrote a higher-order filtering function to filter a dataframe by some function
on its columns. For the following summary of genre, only rows with a recorded
genre are included.

```{r}
# Remove all rows where there is a missing value for col #
filterOutMissingValues = function(df, col)
{
  return(filterBy(df, col, (function(x) {!(is.na(x))})))
}
```

### Genre

In these pie-charts for genre, a single movie can appear multiple times in the
chart. If the movie has tags "Action" and "Comedy" it will be counted twice.

```{r, echo=FALSE}
will.genres = filterOutMissingValues(will.df$movies, 6);
gdf = buildGenresDf(will.genres);
plotGenresDf(gdf, "Will Smith Movies by Genre");
```

```{r, echo=FALSE}
denzel.genres = filterOutMissingValues(denzel.df$movies, 6);
gdf = buildGenresDf(denzel.genres);
plotGenresDf(gdf, "Denzel Washington Movies by Genre");
```

### MPAA Rating

These pie-charts provide a visual for MPAA ratings for Will Smith / Denzel 
Washington movies.

```{r, echo=FALSE}
will.mpaa = filterOutMissingValues(will.df$movies, 7);
mdf = buildMPAADf(will.mpaa);
plotMPAADf(mdf, "Will Smith MPAA Ratings");
```

```{r, echo=FALSE}
denzel.mpaa = filterOutMissingValues(denzel.df$movies, 7);
mdf = buildMPAADf(denzel.mpaa);
plotMPAADf(mdf, "Denzel Washington MPAA Ratings");
```

### Average Audience Rating

The average audience ratings for each actor.

```{r, echo=FALSE}
will.ratings = filterOutMissingValues(will.df$movies, 9);
denzel.ratings = filterOutMissingValues(denzel.df$movies, 9);
rdf <- as.data.frame(matrix(0, ncol = 2, nrow = 2));
names(rdf)[1] <- "actor";
names(rdf)[2] <- "avg.rating";
rdf[1,1] <- "Will";
rdf[2,1] <- "Denzel";
rdf[1,2] <- (sum(will.ratings$ratings)) / nrow(will.ratings);
rdf[2,2] <- (sum(denzel.ratings$ratings)) / nrow(denzel.ratings);
rdf
p <- ggplot(data=rdf, aes(x=actor, y=avg.rating, fill=avg.rating)) 
p <- p +  geom_bar(stat="identity", position=position_dodge())
p <- p + xlab("Actor");
p <- p + ylab("Avg. Rating");
p <- p + theme(legend.position="none")
p
```





















