---
title: 'R Notebook: o-well'
author: "Peyton Urquhart [peyton.urquhart]"
output:
  html_document:
    df_print: paged
---

```{r, echo=FALSE}
require("ggplot2")
which.github = "local";
github.local = "C:/_git_/peyton.urquhart/";
github.remote = "https://raw.githubusercontent.com/peytonurquhart/WSU_STAT_419/main/";

local.metals = paste0( github.local, "datasets/o-well/wells-metals.csv");
local.chem = paste0( github.local, "datasets/o-well/wells-chemistry.csv");
local.wellsRDS = paste0( github.local, "datasets/o-well/wells.rds");

if(which.github == "remote")
{
  include.me = paste0( github.remote, "functions/functions-owell.R");
  library(devtools)
  library(rstudioapi)
  source_url(include.me);
  
  # Download data to working directory if using remotely
  include.metals = paste0( github.remote, "datasets/o-well/wells-metals.csv");
  include.chem = paste0( github.remote, "datasets/o-well/wells-chemistry.csv");
  local.metals = paste0(dirname(getActiveDocumentContext()$path), "/wells-metals.csv")
  local.chem = paste0(dirname(getActiveDocumentContext()$path), "/wells-chemistry.csv");
  local.wellsRDS = paste0(dirname(getActiveDocumentContext()$path), "/wells.RDS");
  download.file(url=include.metals, destfile=local.metals, method='curl');
  download.file(url=include.chem, destfile=local.chem, method='curl');
  
} else if (which.github == "local") {
  
    include.me = paste0( github.local, "functions/functions-owell.R");
    source(include.me);
}

print("Retreived Well Datasets At:");
print(local.metals);
print(local.chem);
#Now, all well csv data it local
```

# Gathering Well Data
Well data is stored in a Comma Separated Values files (.csv), it can then be directly 
read into a dataframe using the read.csv() function.
All necessary data was added to the .csv files prior to reading the data to a dataframe.

# Well Data
owell.metals contains data of dissolved metals detected in water-samples from wells
1-23 in ug/L.

owell.chemistry contains datapoints for physical and chemical paramaters
in ground-water samples from wells 1-23 in mg/L.

Additionally, each dataset includes location and altitude (Feet) for the wells, approximate
distance from the nearest fault-line (Meters) and the surrounding geology.

owell.metals:
```{r}
owell.metals = readWellCsvData(local.metals);
```

owell.chemistry:
```{r}
owell.chemistry = readWellCsvData(local.chem);
```

owell.metals and owell.chemistry are placed into a list to be refrenced by
wells\$saudi\$chemistry and wells\$saudi\$metals
```{r, echo=FALSE}
wells <- list(saudi=NULL);
wells$saudi <- list(chemistry=owell.chemistry, metals=owell.metals);
print(wells);
saveRDS(wells, file = local.wellsRDS);
```

Average metal content in the water for wells 1-23:
```{r, echo=FALSE}
plotMeanMetalsForWellData(wells$saudi$metals);
```

Standard deviation for metal content in the water for wells 1-23:
```{r, echo=FALSE}
plotStDevMetalsForWellData(wells$saudi$metals);
```

Average chemical content in the water for wells 1-23:
```{r, echo=FALSE}
plotMeanChemForWellData(wells$saudi$chemistry);
```

Standard deviation for chemical content in the water for wells 1-23:
```{r, echo=FALSE}
plotStDevChemForWellData(wells$saudi$chemistry);
```


















