---
title: 'R Notebook: o-well'
author: "Peyton Urquhart [peyton.urquhart]"
output:
  html_document:
    df_print: paged
---

```{r, echo=FALSE}
require("ggplot2")
which.github = "local";
github.local = "C:/_git_/peyton.urquhart/";
github.remote = "https://raw.githubusercontent.com/peytonurquhart/WSU_STAT_419/main/";

local.metals = paste0( github.local, "datasets/o-well/wells-metals.csv");
local.chem = paste0( github.local, "datasets/o-well/wells-chemistry.csv");
local.wellsRDS = paste0( github.local, "datasets/o-well/wells.rds");

if(which.github == "remote")
{
  include.me = paste0( github.remote, "functions/functions-owell.R");
  library(devtools)
  library(rstudioapi)
  source_url(include.me);
  
  # Download data to working directory if using remotely
  include.metals = paste0( github.remote, "datasets/o-well/wells-metals.csv");
  include.chem = paste0( github.remote, "datasets/o-well/wells-chemistry.csv");
  local.metals = paste0(dirname(getActiveDocumentContext()$path), "/wells-metals.csv")
  local.chem = paste0(dirname(getActiveDocumentContext()$path), "/wells-chemistry.csv");
  local.wellsRDS = paste0(dirname(getActiveDocumentContext()$path), "/wells.RDS");
  download.file(url=include.metals, destfile=local.metals, method='curl');
  download.file(url=include.chem, destfile=local.chem, method='curl');
  
} else if (which.github == "local") {
  
    include.me = paste0( github.local, "functions/functions-owell.R");
    source(include.me);
}

print("Retreived Well Datasets At:");
print(local.metals);
print(local.chem);
#Now, all well csv data it local
```

# Gathering Well Data
Well data is stored in a Comma Separated Values files (.csv), it can then be directly 
read into a dataframe using the read.csv() function.
All necessary data was added to the .csv files prior to reading the data to a dataframe.

# Well Data
owell.metals contains data of dissolved metals detected in water-samples from wells
1-23 in ug/L.

owell.chemistry contains datapoints for physical and chemical paramaters
in ground-water samples from wells 1-23 in mg/L.

Additionally, each dataset includes location and altitude (Feet) for the wells, 
approximate distance from the nearest fault-line (Meters), approximate distance 
from the saltwater gulf (Meters), and the surrounding geology.

```{r, echo=FALSE}
owell.metals = readWellCsvData(local.metals);
```

```{r, echo=FALSE}
owell.chemistry = readWellCsvData(local.chem);
```

owell.metals and owell.chemistry are placed into a list to be refrenced by
wells\$saudi\$chemistry and wells\$saudi\$metals
```{r, echo=FALSE}
wells <- list(saudi=NULL);
wells$saudi <- list(chemistry=owell.chemistry, metals=owell.metals);
#print(wells);
saveRDS(wells, file = local.wellsRDS);
```

## Mean Water Contents

```{r, echo=FALSE}
plotMeanMetalsForWellData(wells$saudi$metals);
```

```{r, echo=FALSE}
plotMeanChemForWellData(wells$saudi$chemistry);
```

## Coefficients of Variance for Water Contents

```{r, echo=FALSE}
plotCVMetalsForWellData(wells$saudi$metals);
```

```{r, echo=FALSE}
plotCVChemForWellData(wells$saudi$chemistry);
```


# Mercury Content in Well Water

"The U.S. Environmental Protection Agency (EPA) sets a Maximum Contaminant 
Level (MCL) for mercury in drinking water of 0.002 mg/L" (https://epi.dph.ncdhhs.gov/).

In other words, It has been determined that a concentration of Hg greater than
2 micrograms per liter (ug/L) is potentially unsafe for drinking.

From the above data, the mean Hg content for wells 1-23 is 10 ug/L (about than
5 times what is considered safe) and the coefficient of variance is high. This
indicates that there are outside factors which significantly influence concentration
of Hg in the water and may bring it to dangerous levels.



```{r, echo=FALSE}
df = wells$saudi$metals
plotPearsonCorrelation(df, df$dist.gulf, df$Hg, "Distance from Gulf vs Hg Content: Pearson Correlation", "Dist. Gulf (M)", "Hg (ug/L)");
```

There is a weak correlation between a wells distance from the gulf, and
its mercury content. As wells become closer to the gulf, there appears to be a 
greater probability for very high levels of mercury in a water sample.

This can be further explored by examining mercury content based on a wells
sandy/not sandy surrounding geology. Perhaps, water from the gulf seeking into
the source for the well becomes more prominent when the type of soil is sandy
as opposed to rocky.


```{r, echo=FALSE}
plotHgContentVsSand(wells$saudi$metals);
```














