---
title: 'R Notebook: data-intro'
author: "Peyton Urquhart [peyton.urquhart]"
output:
  html_document:
    df_print: paged
---

```{r, echo=FALSE}
require("ggplot2")
```

```{r}
github.local = "C:/_git_/peyton.urquhart/";
github.remote = "https://raw.githubusercontent.com/peytonurquhart/WSU_STAT_419/main/";
script = "functions/functions-imdb.R";

# change this
which.github = "local";
#which.github = "remote";
```

```{r, echo=FALSE}
if(which.github == "remote")
{
  include.me = paste0( github.remote, script);
  library(devtools)
  library(rstudioapi)
  source_url(include.me);
  
} else if (which.github == "local") {
  
    include.me = paste0( github.local, script);
    source(include.me);
}
```

```{r, chunk-setup, include=FALSE}
# include=FALSE means it is running, but invisible ...
knitr::opts_chunk$set(echo = TRUE);
knitr::opts_chunk$set(warning = FALSE);
knitr::opts_chunk$set(message = FALSE);
```

```{r}
#install.packages("C:/_git_/humanVerseWSU/humanVerseWSU_0.1.4.zip", repos=NULL, type="source");
library(humanVerseWSU);

#install.packages("C:/_git_/monte-imdb/imdb_0.1.1.zip", repos=NULL, type="source");
library(imdb);
```

```{r, chunck-loadData, include=TRUE}
loadDataIMDB();
```

## Will Smith

Build a dataframe "will.df" for Will Smith with 2 primary sections:

will.df$info containing a single row of information about Will Smith.

will.df$movies containing a large dataframe where each row represents one of Will
Smiths movies.

```{r}
will.df <- list(movies = IMDB.getMoviesForPerson("nm0000226"), info = IMDB.getPersonInfo("nm0000226"));
will.df$movies;
will.df$info;
```

## Denzel Washington

Build a dataframe "denzel.df" for Denzel Washington with 2 primary sections:

denzel.df$info containing a single row of information about Denzel.

denzel.df$movies containing a large dataframe where each row represents one of
Denzel Washingtons movies.

```{r}
denzel.df <- list(movies = IMDB.getMoviesForPerson("nm0000243"), info = IMDB.getPersonInfo("nm0000243"));
denzel.df$movies;
denzel.df$info;
```

## Common Movies

Search the dataframes for matching ttid(s). Build a new dataframe for each match
resulting in a dataframe of movies starring both Will and Denzel

```{r}
findCommonMovies(will.df$movies, denzel.df$movies);
```


## Summary





































