---
title: 'will-is-better playground'
author: "Peyton Urquhart [peyton.urquhart]"
output:
  html_document:
    df_print: paged
---

```{r, echo=FALSE}
require("ggplot2")
library(humanVerseWSU);
library(imdb);
```

```{r, echo=FALSE}
github.local = "C:/_git_/peyton.urquhart/";
github.remote = "https://raw.githubusercontent.com/peytonurquhart/WSU_STAT_419/main/";
script = "functions/functions-imdb.R";
```

#### Local or Remote ?

```{r}
which.github = "local";
#which.github = "remote";
```

```{r, echo=FALSE}
if(which.github == "remote")
{
  include.me = paste0( github.remote, script);
  library(devtools)
  library(rstudioapi)
  source_url(include.me);
  
} else if (which.github == "local") {
  
    include.me = paste0( github.local, script);
    source(include.me);
}
```

```{r, chunk-setup, include=FALSE, echo=FALSE}
# include=FALSE means it is running, but invisible ...
knitr::opts_chunk$set(echo = TRUE);
knitr::opts_chunk$set(warning = FALSE);
knitr::opts_chunk$set(message = FALSE);
```

```{r, chunck-loadData, include=TRUE}
loadDataIMDB();
```

## Will Smith

```{r}
will.df <- list(movies = IMDB.getMoviesForPerson("nm0000226"), info = IMDB.getPersonInfo("nm0000226"));
will.df$movies;
will.df$info;
```

## Denzel Washington

```{r}
denzel.df <- list(movies = IMDB.getMoviesForPerson("nm0000243"), info = IMDB.getPersonInfo("nm0000243"));
denzel.df$movies;
denzel.df$info;
```

## Average Audience Rating

The overall average audience ratings for each actor.

```{r, echo=FALSE}
plotAvgAudienceRating(will.df, denzel.df);
```


### In Comedy

```{r, echo=FALSE}

will.new = filterBy(filterOutMissingValues(will.df$movies, 6), 6, (function(x) {grepl("Comedy", x)}))
denzel.new = filterBy(filterOutMissingValues(denzel.df$movies, 6), 6, (function(x) {grepl("Comedy", x)}))
                       
plotAvgAudienceRating(list(movies = will.new), list(movies = denzel.new))
```

### In Drama

```{r, echo=FALSE}

will.new = filterBy(filterOutMissingValues(will.df$movies, 6), 6, (function(x) {grepl("Drama", x)}))
denzel.new = filterBy(filterOutMissingValues(denzel.df$movies, 6), 6, (function(x) {grepl("Drama", x)}))
                       
plotAvgAudienceRating(list(movies = will.new), list(movies = denzel.new))
```

### In Action

```{r, echo=FALSE}

will.new = filterBy(filterOutMissingValues(will.df$movies, 6), 6, (function(x) {grepl("Action", x)}))
denzel.new = filterBy(filterOutMissingValues(denzel.df$movies, 6), 6, (function(x) {grepl("Action", x)}))
                       
plotAvgAudienceRating(list(movies = will.new), list(movies = denzel.new))
```

### In Adventure

```{r, echo=FALSE}

will.new = filterBy(filterOutMissingValues(will.df$movies, 6), 6, (function(x) {grepl("Adventure", x)}))
denzel.new = filterBy(filterOutMissingValues(denzel.df$movies, 6), 6, (function(x) {grepl("Adventure", x)}))
                       
plotAvgAudienceRating(list(movies = will.new), list(movies = denzel.new))
```

### In Romance

```{r, echo=FALSE}

will.new = filterBy(filterOutMissingValues(will.df$movies, 6), 6, (function(x) {grepl("Romance", x)}))
denzel.new = filterBy(filterOutMissingValues(denzel.df$movies, 6), 6, (function(x) {grepl("Romance", x)}))
                       
plotAvgAudienceRating(list(movies = will.new), list(movies = denzel.new))
```

### In Crime

```{r, echo=FALSE}

will.new = filterBy(filterOutMissingValues(will.df$movies, 6), 6, (function(x) {grepl("Crime", x)}))
denzel.new = filterBy(filterOutMissingValues(denzel.df$movies, 6), 6, (function(x) {grepl("Crime", x)}))
                       
plotAvgAudienceRating(list(movies = will.new), list(movies = denzel.new))
```

### In Sport

```{r, echo=FALSE}

will.new = filterBy(filterOutMissingValues(will.df$movies, 6), 6, (function(x) {grepl("Sport", x)}))
denzel.new = filterBy(filterOutMissingValues(denzel.df$movies, 6), 6, (function(x) {grepl("Sport", x)}))
                       
plotAvgAudienceRating(list(movies = will.new), list(movies = denzel.new))
```

It turns out, Denzel wins in every single category...


## Total Votes

Each user on IMDB may "vote" on a movie. A vote indicates they are interested-in 
and have watched the movie.

```{r}
plotAvgVotes(will.df, denzel.df);
```

### In Comedy

```{r, echo=FALSE}

will.new = filterBy(filterOutMissingValues(will.df$movies, 6), 6, (function(x) {grepl("Comedy", x)}))
denzel.new = filterBy(filterOutMissingValues(denzel.df$movies, 6), 6, (function(x) {grepl("Comedy", x)}))
                       
plotAvgVotes(list(movies = will.new), list(movies = denzel.new))
```

### In Action

```{r, echo=FALSE}

will.new = filterBy(filterOutMissingValues(will.df$movies, 6), 6, (function(x) {grepl("Action", x)}))
denzel.new = filterBy(filterOutMissingValues(denzel.df$movies, 6), 6, (function(x) {grepl("Action", x)}))
                       
plotAvgVotes(list(movies = will.new), list(movies = denzel.new))
```

### In Drama

```{r, echo=FALSE}

will.new = filterBy(filterOutMissingValues(will.df$movies, 6), 6, (function(x) {grepl("Drama", x)}))
denzel.new = filterBy(filterOutMissingValues(denzel.df$movies, 6), 6, (function(x) {grepl("Drama", x)}))
                       
plotAvgVotes(list(movies = will.new), list(movies = denzel.new))
```

## Box-office profit

### Top 10 Movies

```{r, echo=FALSE}

will.new = filterBy(filterOutMissingValues(will.df$movies, 3), 3, (function(x) {x <= 10}))
will.new
                       
plotBoxOffice(list(movies = will.new), list(movies = denzel.new))
```

























