---
title: 'R Notebook: r-intro'
author: "Peyton Urquhart [peyton.urquhart]"
output:
  html_document:
    df_print: paged
---

```{r, echo=FALSE}
require("ggplot2")
```


```{r, echo=FALSE}
which.github = "local";
github.local = "C:/_git_/peyton.urquhart/";
github.remote = "https://raw.githubusercontent.com/peytonurquhart/WSU_STAT_419/main/";

local.draft = paste0( github.local, "datasets/declaration/draft.txt");
local.final = paste0( github.local, "datasets/declaration/final.txt");

if(which.github == "remote")
{
  include.me = paste0( github.remote, "functions/functions-intro.R");
  library(devtools)
  library(rstudioapi)
  source_url(include.me);
  
  # Download data to working directory if using remotely
  include.draft = paste0( github.remote, "datasets/declaration/draft.txt");
  include.final = paste0( github.remote, "datasets/declaration/final.txt");
  local.draft = paste0(dirname(getActiveDocumentContext()$path), "/draft.txt")
  local.final = paste0(dirname(getActiveDocumentContext()$path), "/final.txt");
  download.file(url=include.draft, destfile=local.draft, method='curl');
  download.file(url=include.final, destfile=local.final, method='curl');
  
} else if (which.github == "local") {
  
    include.me = paste0( github.local, "functions/functions-intro.R");
    source(include.me);
}

print("Retreived Declaration of Independence at:");
print(local.draft);
print(local.final);
#Now, all declaration drafts are local

```


# Handshake
handShake(n, plotMe) calculates the total number of handshakes that must occur
between n people such that each individual shakes everybody's hand exactly one time.
Handshakes may be visualized using plotMe=TRUE.

Handshake plot for n = 4 people:
```{r, echo=FALSE}
handShake(n=4, plotMe=TRUE)
```

Handshake plot for n = 8 people:
```{r, echo=FALSE}
handShake(n=8, plotMe=TRUE)

```

Handshake plot for n = 24 people:
```{r, echo=FALSE}
handShake(n=24, plotMe=TRUE)
```

## Function Parameters
*  n : Total number of people that must shake everybody's hand exactly once

*  plotMe : Print a plot of the handshake simulation?

## Calculation of plotMe
circleData (c(x,y), r, npoints):
Generates data consisting of n points, each at equal and increasing angles 
from c (the center) which (when connected) form a circle. This is achieved by
generating a sequence s of radians (0, ... ,2*PI) in equal intervals, followed
by two more sequences x,y where x[i] = (c.x + r cos(s[i])) and y[i] = (c.y + r sin(s[i])).

circleData() is called once to plot an initial circle drawn with lines, and once again with
npoints = n to plot individual points corresponding to each person in the handshake simulation.

Finally, for each individual point on the circle, n - 1 lines are drawn connecting
the given point with all other points on the circle.


# Alphabet Declaration of Independence
countLowerCaseChars_f(filepath): Counts lower case alphabetical characters in a .txt file and
returns a data frame. Here it is used to count chars in the Declaration of Independence
draft and final copies.

Declaration of independence draft table:
```{r, echo=FALSE}
countLowerCaseChars_f(local.draft);
```

Declaration of independence final table:
```{r, echo=FALSE}
countLowerCaseChars_f(local.final);
```


```{r, echo=FALSE}
df1 = countLowerCaseChars_f(local.draft);
df2 = countLowerCaseChars_f(local.final);
plotHistogramForAlphabetData(df1, df2);
```


# Determinant of 3x3 matrix

$3 \times 3$


